---
title: "homework3_cq2207"
author: "Carolina Q Cardoso"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r problem1}



library(tidyverse)
library(dplyr)
library(ggplot2)
library(viridis)
library(ggridges) 


library(p8105.datasets)
data("instacart")


```
The dataset `Instacart` contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. Variables describe details of the order and of the products ordered.

```{r problem1_cont}

#How many aisles are there, and which aisles are the most items ordered from?

aisle_count = instacart %>%
  group_by(aisle_id, aisle) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

```

There are `r nrow(aisle_count)` different aisles. The aisle with the highest number of ordered items in "`r pull(aisle_count[1,2])`" with `r pull(aisle_count[1,3])` items, followed by "`r pull(aisle_count[2,2])`" with `r pull(aisle_count[2,3])` ordered items.


```{r - problem1_cont2}
#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

library(ggplot2)

aisle_count %>%
  filter(count > 10000) %>%
  ggplot(aes(x = aisle_id, y = count)) + 
    geom_point() + geom_line() + 
    labs(
    title = "Number of Items Ordered by Aisle (n > 10000)",
    x = "Aisle",
    y = "Total Number of Items Ordered")

#Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

library(knitr)

product_count = instacart %>%
  group_by(aisle, product_name) %>%
  summarize (count = n()) %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  mutate (rank = min_rank(desc(count))) %>%
  filter (rank <= 3) %>%
  select(-rank) %>% 
  knitr::kable() %>% 
    print()
 

#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
  

library(tidyverse)
library(tidyr)

apple_coffee = instacart %>% 
  group_by (product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream")%>%
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>%
  knitr::kable(col.names = c("Product", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")) %>% 
    print()

```


```{r problem2}

data("brfss_smart2010")

brfss_new = brfss_smart2010 %>% 
    janitor::clean_names() %>%
    mutate_all(tolower) %>%
    filter (topic == "overall health") %>%
    rename(`state` = locationabbr, `county` = locationdesc, `coordinates` = geo_location) %>%
    mutate(response = forcats::fct_relevel(response, c("poor","fair","good","very good","excellent")))

  
brfss_new 

 #In 2002, which states were observed at 7 or more locations? What about in 2010?

brfss_2002 = brfss_new %>%
    group_by(year, state) %>%
    summarize(n = n_distinct(coordinates)) %>%
    filter (n > 6, year == '2002')

brfss_2010 = brfss_new %>%
    group_by(year, state) %>%
    summarize(n = n_distinct(coordinates)) %>%
    filter (n > 6, year == '2010')
  
```

In 2002, there were `r nrow(brfss_2002)` states that were observed in 7 or more locations: `r pull(brfss_2002[1,2])`, `r pull(brfss_2002[2,2])`, `r pull(brfss_2002[3,2])`, `r pull(brfss_2002[4,2])`, `r pull(brfss_2002[5,2])`, and `r pull(brfss_2002[6,2])`.
In 2010, there were `r nrow(brfss_2010)` states that were observed in 7 or more locations: `r pull(brfss_2010[1,2])`, `r pull(brfss_2010[2,2])`, `r pull(brfss_2010[3,2])`, `r pull(brfss_2010[4,2])`, `r pull(brfss_2010[5,2])`, `r pull(brfss_2010[6,2])`, `r pull(brfss_2010[7,2])`, `r pull(brfss_2010[8,2])`, `r pull(brfss_2010[9,2])`, `r pull(brfss_2010[10,2])`, `r pull(brfss_2010[11,2])`, `r pull(brfss_2010[12,2])`, `r pull(brfss_2010[13,2])`, and `r pull(brfss_2010[14,2])`.

```{r problem2_cont}

#Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

brfss_excellent = brfss_new %>%
  transform(data_value = as.numeric(data_value))%>%
    filter (response == 'excellent') %>%
    group_by(state, year, response) %>%
    summarize(mean_data = mean(data_value))

brfss_excellent %>% 
    ggplot(aes(x = year, y = mean_data, group = state, color = state)) + 
    geom_point() + geom_line() +
    labs(
    title = "Average value over time by state",
    x = "Year",
    y = "Average Data Value"
  )

#Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

brfss_two = brfss_new %>%
    transform(data_value = as.numeric(data_value)) %>%
    filter(state == 'ny',year == '2006' | year == '2010') %>%
    select(state, year, response, data_value)

brfss_two %>%
    ggplot(aes(x = response, y = data_value, color = response)) + 
    geom_boxplot() + 
    facet_grid(. ~ year) + 
    labs(
    title = "Distribution of Data Values per Response in NY State",
    x = "Response",
    y = "Data Value")
    

```

```{r problem3}

#Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

accel_data = read_csv("./data/accel_data.csv") 

accel_new = accel_data %>%
    janitor::clean_names() %>%
    mutate(day = str_to_lower(day)) %>%
    mutate_if(is.numeric, round, digits = 2) %>%
    rename (`day_of_week` = day) %>%
    mutate(weekend = if_else(day_of_week=="sun" | day_of_week == 'sat', "weekend", "weekday")) %>%
    mutate(day_of_week = forcats::fct_relevel(day_of_week, c("mon","tues","wed","thurs","fri", "sat", "sun"))) %>%
    select(week, day_id, day_of_week, weekend, everything()) 

```

The dataset `accel_new` has `r nrow(accel_new)` observations and `r ncol(accel_new)` variables. Variables include the `week` and `day_of_week` when the accelerometer data was collected. The variable `weekend` indicates whether the day the data was collected was a weekend or not. The variable activity_x, from `activity_1` to `activity_1440` represent the activity count for each minute of the day, starting at midnight. 
The dataset shows that this patient wearing the accelerometer used it for `r pull(accel_new[35,1])` weeks and `r pull(accel_new[35,2])` days.

```{r problem3_cont}

#Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

accel_total_activity = accel_new %>%
  pivot_longer(activity_1:activity_1440, names_to = "minute", values_to = "activity_min") %>%
  separate(minute, into = c(NA, "minute"), sep = "activity_") %>%
  mutate(activity_min = as.numeric(activity_min),minute = as.numeric(minute), day_id = as.numeric(day_id)) %>%
  group_by(day_id, day_of_week) %>%
  summarize(activity_day = sum(activity_min)) %>%
  arrange(day_id) 

accel_total_activity %>% knitr::kable() %>% 
    print()

```

Based on the table, there are no apparent trends in the activity level of the patient per day.

```{r problem3_cont2}

#Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

accel_total_activity %>%
    ggplot(aes(x = day_id, y = activity_day, color = day_of_week)) + 
    geom_point() + geom_line() +
    labs(
    title = "Activity Time by Day",
    x = "Day",
    y = "Activity Time")
```

The graph shows that the patient's weekend activity stayed somewhat stable and moderate throught out the entire testing period. The activity level varies by weekday, showing no apparent trends.
