---
title: "homework3_cq2207"
author: "Carolina Q Cardoso"
date: "10/9/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r problem1}

library(tidyverse)

library(p8105.datasets)
data("instacart")


```
The dataset `Instacart` contains `r nrow(instacart)` observations and `r ncol(instacart)` variables. Variables describe details of the order and of the products ordered.

```{r problem1_cont}

#How many aisles are there, and which aisles are the most items ordered from?

aisle_count = instacart %>%
  group_by(aisle_id, aisle) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

```

There are `r nrow(aisle_count)` different aisles. The aisle with the highest number of ordered items in "`r pull(aisle_count[1,2])`" with `r pull(aisle_count[1,3])` items, followed by "`r pull(aisle_count[2,2])`" with `r pull(aisle_count[2,3])` ordered items.


```{r - problem1_cont2}
#Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

library(ggplot2)

aisle_count %>%
  filter(count > 10000) %>%
  ggplot(aes(x = aisle_id, y = count)) + 
    geom_point() + geom_line() + 
    labs(
    title = "Number of Items Ordered by Aisle (n > 10000)",
    x = "Aisle",
    y = "Total Number of Items Ordered")

#Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

library(knitr)

product_count = instacart %>%
  group_by(aisle, product_name) %>%
  summarize (count = n()) %>%
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>%
  mutate (rank = min_rank(desc(count))) %>%
  filter (rank <= 3) %>%
  select(-rank) %>% 
  knitr::kable() %>% 
    print()
 

#Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
  

library(tidyverse)
library(tidyr)

apple_coffee = instacart %>% 
  group_by (product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream")%>%
  pivot_wider(names_from = order_dow, values_from = mean_hour) %>%
  knitr::kable(col.names = c("Product", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")) %>% 
    print()

```